---
title: "GunViolence data"
author: "Keeno Glanville"
date: "2022-10-09"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(tidyr)
library(dplyr)


```

# Prompt:What factors most influences crime? and what state has the largest crime rate?
### Load data

```{r}
set.seed(1324)
urldata <- "https://raw.githubusercontent.com/kglan/MSDS/main/DATA607/Data%20Transformation/GunViolence/guns.csv"
gundata.r <- read_csv(url(urldata))



```



## We first start by cleaning the data set and renaming columns to make the names clearer
```{r}
gundata_clean <- gundata.r %>%
  select(-"...1", -"male")%>%
  rename("%Black" = "afam",
         "%White"= "cauc",
         "Crime/100k" = "violent",
         "Murder/100k" = "murder",
         "Robbery/100k" = "robbery",
         "Prisoners/100k" = "prisoners",
         "Income/capitaofstate" = "income",
         "Density(sqrmi/1000)" = "density",
          ShallCarryLaw = "law",
         "population/mil" = population)
gundata_clean

summary(gundata_clean)

```


### We wil now pair the data together to recognize the trends within our data

```{r}

gundata_pairs <- gundata_clean%>%
  mutate(ShallCarryLaw = ifelse(ShallCarryLaw == "no",0,1))
pairs(data=gundata_pairs,
      ~ year+ `Crime/100k`+ `Murder/100k` + `Robbery/100k`+ `Prisoners/100k`+
        `Density(sqrmi/1000)` + `Income/capitaofstate`)

```

## In the before step we see tremendous trends within the data and now we want to see our correlation coefficients
### We will print corrplot and histograms as well as boxplots

```{r}
library(corrplot)
library(PerformanceAnalytics)

gundata_corplot <- gundata_clean %>%
  select(-"state", -"ShallCarryLaw")

chart.Correlation(gundata_corplot,
                  method="pearson",
                  histogram=TRUE,
                  pch=16)

gundata_cor <- cor(gundata_corplot, method = "pearson")
col_gd <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(gundata_cor, method = "color", col = col_gd(200),
         type = "upper", order = "hclust",
         addCoef.col = "black",
         tl.col = "black", tl.srt = 45,)

```



# BoxPlot shows the trend of crime over the years, with a slight trend upwards heavily skewed by District of Columbia
## (district of columbia heavily reflects the data points at the top of the box plots)

```{r}
library(ggplot2)
ggplot(gundata_clean, aes(x= year, y= `Crime/100k`, group = `year`)) +
         geom_boxplot( fill="skyblue", notch=FALSE) +
         geom_jitter( size=1, color="orange", width=0.3)

```




# Wide data Transformed
## We will now transform the data to get a representation of state vs Crime rate

```{r}

gunstate<- gundata_clean%>%
  select(1,2,11)
gunstate
```



## Now we want to look at plots of crime with regards to state. We see District of Columbia elevated overall
```{r}
library(lattice)
xyplot(log(`Crime/100k`) ~ as.numeric(as.character(year)) | state, data = gunstate, type = "l")

```



# We conclude that we see the data is heavily skewed due to the District of Columbia
## This data set is demonstrating to us that not just one factor influences the crime rate within a given state. We observed the data was heavily skewed for density with the district of Columbia having such a heavily dense population.








