---
title: "Working with Tidy Data"
author: "Keeno Glanville"
date: "2022-10-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
```

## load CSV file and Clean data


```{r}



url <- "https://raw.githubusercontent.com/kglan/MSDS/main/DATA607/WorkingWithTidyData/WorkingwithTidyData.csv"
rawdata <-read.csv(url)
rawdata <- t(rawdata)
rawdata<- data.frame(rawdata)
rownames(rawdata)<- c()
rawdata
names(rawdata)
rownames(rawdata)
data <- rawdata%>%
  select(-3,)%>%
  slice(-c(1,2))%>%
  mutate_if(is.character, as.numeric)%>%
  t()%>%
  data.frame()%>%
  mutate("Total" = X1 + X2 +X3 + X4 + X5)%>%
  t()%>%
  data.frame()%>%
  rename("ALASKA on Time" = "X1",
         "ALASKA delayed" = "X2",
         "AM_WEST on Time" = "X4",
         "AM_WEST delayed "= "X5")

row.names(data) <- c("Los Angeles", "Phoenix", "San Diego", "San Francisco", "Seattle", "Total")
data

```

## Now that data is clean we will conduct Analysis 
When we compare both airlines we see that the ALASKA airline has fewer delays
within their flights. It is significant to note that The ALASKA airline has less
flights total which shows that with such a magnitude of flights AM_WEST is still
close in comparison.

```{r}

airline<- data%>%
  mutate("ALASKA total_flights"= `ALASKA delayed`+`ALASKA on Time`)%>%
  mutate("AM_WEST total_flights"= `AM_WEST delayed `+ `AM_WEST on Time`)%>%
  mutate("ALASKA total_delays"= `ALASKA delayed`/(`ALASKA on Time`+`ALASKA delayed`))%>%
  mutate("AM_WEST total_delays"= `AM_WEST delayed `/(`AM_WEST on Time`+`AM_WEST delayed `))%>%
  round(digits = 2)

airline
           
```



# Visualization
```{r}
library(ggplot2)
airline$City<- c("Los Angeles","Phoenix", "San Diego", "San Francisco", "Seattle", "Total")
ggplot(airline, aes(x=City, y=`ALASKA total_delays`)) +
  geom_bar(position='dodge', stat='identity')
ggplot(airline, aes(x=City, y= `AM_WEST total_delays`)) +
  geom_bar(position=position_dodge(), stat='identity')
```

# Conclusion

When looking at the data in in the bar graphs, we see that AM_WEST has a high
delay rate in San Francisco and Seattle which skew its delay rate with these two
cities. When looking at ALASKA it shows that the delay rates are relatively 
similar amongst the cities besides Phoenix and San Diego. This implies that 
AM_WEST needs to focus on fixing the reasons it has such high delays in San 
Francisco and Seattle.
