---
title: 'Assignment 1: Data Subsetting'
author: "Keeno Glanville"
date: "2022-09-04"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidymodels)
library(readr)

```


# Introduction
I am opting to go with the FIFA data set. https://fivethirtyeight.com/features/how-to-break-fifa/
I chose this data set as it gives insightful info as to the participation of countries to the
FIFA world cup viewership. This is insightful as it demonstrates the strength of viewership in
each area and could be used to group FIFA relations and outreach by continents.
The Business Problem I want to take is Outreach to areas that do not watch the World Cup. FIFA
officials were indicted for bribery of the rights to host the world cup competitions.


## Reading dataset into R
```{r cars}
urldata="https://raw.githubusercontent.com/kglan/MSDS/main/DATA607/1%20Data%20Subsetting/FIFAData"
FIFAdata<-read_csv(url(urldata))
names(FIFAdata)
```

## We will now Subset the data to a new data frame which illustrates the Data of Region with its relation to Global Population and FIFA Viewership


```{r}
FIFAdata <- FIFAdata%>%
  select(-"gdp_weighted_share", -"country")%>%
  rename(Region = "confederation",
         "Global_Population(%)" = "population_share",
         "FIFA_Total_Viewer(%)"= "tv_audience_share")
```

Here we have a dataset which tells us the Region as well as the percentage of
individuals from each country in that region watch the World Cup


# New Data Frame consisting of Region 

```{r}


FIFAtargets <- aggregate(cbind(`Global_Population(%)`,`FIFA_Total_Viewer(%)`) ~ Region,
                             data = FIFAdata,
                             FUN = sum)
FIFAtargets
```


# Conclusion
Seeing as FIFA gave the right to QATAR to host the 2022 FIFA World Cup, we could
make the argument that they are aiming to target regions whose FIFA viewage 
percentage is lower than the Global Population percentage.This could be seen as
CAF and AFC has the only negative ratio.
The counter-argument to their indictment is that they could have given that right
to any country within the AFC region as they have the largest Global Population
as well as the highest percentage of viewership for FIFA. This is all whilst 
having a negative viewership ratio

