---
title: "DATA 606 Data Project Proposal"
author: "Keeno Glanville"
date: "2022-10-23"
output: html_document
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data
library(ggplot2)
library(corrplot)
library(manifestoR)
library(readr)
library(dplyr)
url.test<-"https://raw.githubusercontent.com/kglan/MSDS/main/DATA606/DATA%20606%20Data%20Project%20Proposal/test.csv"
testdata <- read_csv(url(url.test))


url.train <- "https://raw.githubusercontent.com/kglan/MSDS/main/DATA606/DATA%20606%20Data%20Project%20Proposal/train.csv"
traindata <- read_csv(url(url.train))

df<- bind_rows(traindata, testdata)

#Categorical Vairables
cat_features <- df%>%
  select(5,6,8,9,10,12,13,14,16,17,18,20,21)%>%
  colnames()%>%
  as.vector()%>%
  append("bankruptcy")
cat_features

#Numerical Variables
num_features <-df%>%
  select(2,3,4,7,11,15,19)%>%
  colnames()%>%
  as.vector()
num_features

corrro <-df%>%
  select(2,3,4,7,11,15,19) 
```
```{r}
df
```

### Research question 
 Using a Logistic Regression, what is the likelihood that the individuals within the testdata set will have a bankruptcy?



### Cases 

Each case represents individuals who have utilized American Express as their financial institution. There are 1000 observations



### Data collection 

The Data was Collected by American Express as a part of their coding competition hosted on Kaggle.com. The Data is utilized as a method of improving analytical skills



### Type of study 
 This is an observational study



### Data Source 

https://www.kaggle.com/competitions/bankruptcy-risk-prediction/data?select=train.csv



### Dependent Variable

The response variable is bankruptcy, it is binary categorical



### Independent Variable(s)
There are many explanatory variables which contribute to this dataset as many factors could contribute. The magnitude of the factors which contribute will be explored in the Analysis and Logistic Regression model. 



### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
summary(df)

ggplot(df, aes(x=sum)) + 
  geom_histogram()

#Demonstrate correlation between numerical values
library(corrplot)
Tracefile_cor <- cor(corrro, method = "pearson")
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(Tracefile_cor, method = "color", col = col(200),
         type = "upper", order = "hclust",
         addCoef.col = "black",
         tl.col = "black", tl.srt = 45,)


      

```
