---
title: "Tidy-text-mining - John F Kennedy's Inaugural Address"
author: "Keeno Glanville"
date: "2022-11-07"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Data
I tried forking the repository to get the data however when I ran the codes they were very ahrd to run as the dependencies seem outdate4d so I essentially copied the main functions needed

Reference for expenditure:  John F Kennedy's Inaugural Address : https://www.gutenberg.org/cache/epub/3/pg3-images.html
Reference for repository: https://github.com/dgrtwo/tidy-text-mining/blob/5da7251f492c365f7b7e994ac3b5d642e7240194/02-sentiment-analysis.Rmd

```{r}
library(tidytext)
library(janeaustenr)
library(dplyr)
library(stringr)
library(wordcloud)
```

```{r}
get_sentiments("bing")
get_sentiments("afinn")
get_sentiments("nrc")




tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
austen_books()

nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")
nrc_joy

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)


library(tidyr)
jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
jane_austen_sentiment

library(ggplot2)
ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```
## 2.2 Comparing the three sentiment dictionaries

```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")
pride_prejudice

afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")
afinn

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
bing_and_nrc

bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")

get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)

get_sentiments("bing") %>% 
  count(sentiment)
```

## 2.4 Most common positive and negative words 

```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()
bing_word_counts

bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)

custom_stop_words <- bind_rows(tibble(word = c("miss"),  
                                      lexicon = c("custom")), 
                               stop_words)
custom_stop_words
```

## 2.5 Wordclouds

```{r}
library(wordcloud)
tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))

library(reshape2)
tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```


## 2.6 Looking at units beyond just words 

```{r}
p_and_p_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")

p_and_p_sentences$sentence[2]

austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())

bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")
wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())
tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup()
```

# John F. Kennedy's Inaugural Address (extend)

I decided to utilize the gutenberg library
```{r}
library("gutenbergr")
gutenberg_works() %>%
  filter(title == "John F. Kennedy's Inaugural Address")
```
Inaugural Address
```{r}
JFK <- gutenberg_download(3)

text <- tibble(line = 1:nrow(JFK), JFK$text)
colnames(text) <- c('lines', 'text')

clean_book <- text %>%
  unnest_tokens(word, text) #splits a columns into tokens

#Positives
clean_book %>% 
  inner_join(get_sentiments("bing")) %>%
  count(word, sort = TRUE)
  
``` 


## World Clouds
```{r}
#Positive Wordcloud
clean_book %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

```{r}
#Negative Word Cloud
clean_book %>%
  inner_join(get_sentiments("nrc")) %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

```{r}
bing_word_counts <- clean_book %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort=TRUE)

bing_word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ggplot(aes(reorder(word, n), n, fill = sentiment)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment", x = NULL) +
  coord_flip()
```
# Conclusion

John F Kennedy's Inaugural Address was one that promoted and fostered peace and freedom within the nature. He ensured to express his concern for the well being of the nation's constituents. Conversely, he also alluded to the ideas of poverty and tyranny. From this sentiment analysis my personal elicitation without reading the actual article is that JFK was hoping to inspire hope for the future to come of the USA.







