---
title: "Discussion12"
author: "Keeno Glanville"
date: "2023-04-21"
output: html_document
---

```{r setup, include=FALSE}
set.seed(1324)
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
```

# Determine if the average price of a house in NY is decreasing or increasing compared to one year ago.  Will involve turning the months/years into a long format.

## Load data
```{r}
urldat <- "https://raw.githubusercontent.com/kglan/MSDS/main/DATA605/Regression%20Analysis/CarPrice_Assignment.csv"
zill<- read_csv(url(urldat))
zill
```


Multiple Regression



```{r}
model <- lm(price~., data =zill)
summary(model)


```






```{r}
plot(model)



```

# Change terms for binomial and quadratic

```{r cars}

zill$fueltype <- ifelse(zill$fueltype == "gas", 1, 0)
zill$LxW <- zill$carlength * zill$carwidth
zill2<-zill%>%
  select(-c(carlength, carwidth))

```

Rerun model with additions

```{r}

model2 <- lm(price~., data =zill2)
summary(model2)


```



```{r}
plot(model2)



```




# Conclusion 
The second plot including the binomial and the quadratic is actually less accurate and the Q-Q plot shows the data is not distributed normally, however we have strong 
R2 values. I would think with some normalization we could eventually utilize a strong linear model.













