---
title: "Zillow Single Family Home Prices"
author: "Keeno Glanville"
date: "2022-10-09"
output: html_document
---

```{r setup, include=FALSE}
set.seed(1324)
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
```

# Determine if the average price of a house in NY is decreasing or increasing compared to one year ago.  Will involve turning the months/years into a long format.

## Load data
```{r}
urldat <- "https://raw.githubusercontent.com/kglan/MSDS/main/DATA607/Data%20Transformation/ZillowHomePrices/ZillowHomeprices.csv"
zill<- read_csv(url(urldat))
zill
```

# Wide data brought to Long
## Clean and downsize data
( A problem arose where some of the data had NA values, those rows without accurate and conssiten reporting data were removed)
```{r}

zill <- filter(zill, State == "NY")
zill<- zill%>%
  na.omit()%>%
  select(c(250:281))



#Find sum on columns
price<-colMeans(zill)

#Establish Long data view
zillowNY <- data.frame(price)
zillowNY$Date <- row.names(zillowNY)
rownames(zillowNY) <- 1:nrow(zillowNY)


zillowNY 

```

# Now we do a ggplot to see the trend of House Prices over the year



```{r}
ggplot(zillowNY, aes(x=Date, y= price))+
  geom_point()+
  geom_line()+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

# Conclusion

## We see that house prices in New York have been strongly rising from the previous year to now.
