---
title: 'Inference for numerical data'
author: "Keeno Glanville"
output:
  pdf_document: default
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
```

## Getting Started

### Load packages

In this lab, we will explore and visualize the data using the **tidyverse** suite of packages, and perform statistical inference using **infer**. The data can be found in the companion package for OpenIntro resources, **openintro**.

Let's load the packages.

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(infer)
```


### The data

Every two years, the Centers for Disease Control and Prevention conduct the Youth Risk Behavior Surveillance System (YRBSS) survey, where it takes data from high schoolers (9th through 12th grade), to analyze health patterns. You will work with a selected group of variables from a random sample of observations during one of the years the YRBSS was conducted.

Load the `yrbss` data set into your workspace.

```{r load-data}
data('yrbss', package='openintro')
```

There are observations on 13 different variables, some categorical and some numerical. The meaning of each variable can be found by bringing up the help file:

```{r help-nc, eval=FALSE}
?yrbss
```


1.  What are the cases in this data set? How many cases are there in our sample?
13,583

Remember that you can answer this question by viewing the data in the data viewer or by using the following command:

```{r str}
glimpse(yrbss)
```

## Exploratory data analysis

You will first start with analyzing the weight of the participants in kilograms: `weight`.

Using visualization and summary statistics, describe the distribution of weights. The `summary` function can be useful.

```{r summary}
summary(yrbss$weight)
```

2.  How many observations are we missing weights from?
 1004 

Next, consider the possible relationship between a high schooler's weight and their physical activity. Plotting the data is a useful first step because it helps us quickly visualize trends, identify strong associations, and develop research questions.

First, let's create a new variable `physical_3plus`, which will be coded as either "yes" if they are physically active for at least 3 days a week, and "no" if not.


```{r create new var}
yrbss <- yrbss %>% 
  mutate(physical_3plus = ifelse(yrbss$physically_active_7d > 2, "yes", "no"))
```


3.  Make a side-by-side boxplot of `physical_3plus` and `weight`. Is there a relationship between these two variables? What did you expect and why?
 It generally shows a larger outline for those who are not physically active but it is somewhat similar to those that are physically active. I did not expect that. I expected the one's that excercised to be less weight
```{r}
boxplot(yrbss$weight ~ yrbss$physical_3plus, 
        main="Weight and Physical 3 Plus",
        ylab="Weight", 
        xlab="Phy Active 3+ Days", 
        horizontal = FALSE)
```


The box plots show how the medians of the two distributions compare, but we can also compare the means of the distributions using the following to first group the data by the `physical_3plus` variable, and then calculate the mean `weight` in these groups using the `mean` function while ignoring missing values by setting the `na.rm` argument to `TRUE`.

```{r by-means}
yrbss %>%
  group_by(physical_3plus) %>%
  summarise(mean_weight = mean(weight, na.rm = TRUE))
```

There is an observed difference, but is this difference statistically significant? In order to answer this question we will conduct a hypothesis test.

## Inference

4.  Are all conditions necessary for inference satisfied? Comment on each. You can compute the group sizes with the `summarize` command above by defining a new variable with the definition `n()`.
Yes, independence within groups, independence between groups and the size of the each group is sufficiently large
```{r }
yrbss %>%
  group_by(physical_3plus) %>%
  summarise(mean_weight = mean(weight, na.rm = TRUE), count = n())
```
5.  Write the hypotheses for testing if the average weights are different for those who exercise at least times a week and those who don't.

u1 = mean weight of those students physically active >= least 3 days a week
u2 = mean weight of those students physically active < than 3 days a week

H0:u1=u2
HA: u1 != u2


Next, we will introduce a new function, `hypothesize`, that falls into the `infer` workflow. You will use this method for conducting hypothesis tests. 

But first, we need to initialize the test, which we will save as `obs_diff`.

```{r inf-weight-habit-ht-initial, tidy=FALSE, warning = FALSE}

obs_diff <- yrbss %>%
  filter(! is.na(weight)) %>%
  filter(! is.na(physical_3plus)) %>%
  specify(weight ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
```

Notice how you can use the functions `specify` and `calculate` again like you did for calculating confidence intervals. Here, though, the statistic you are searching for is the difference in means, with the order being `yes - no != 0`.

After you have initialized the test, you need to simulate the test on the null distribution, which we will save as `null`.

```{r inf-weight-habit-ht-null, tidy=FALSE, warning = FALSE}
null_dist <- yrbss %>%
  filter(! is.na(weight)) %>%
  filter(! is.na(physical_3plus)) %>%
  specify(weight ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
```

Here, `hypothesize` is used to set the null hypothesis as a test for independence. In one sample cases, the `null` argument can be set to "point" to test a hypothesis relative to a point estimate.

Also, note that the `type` argument within `generate` is set to `permute`, whichis the argument when generating a null distribution for a hypothesis test.

We can visualize this null distribution with the following code:

```{r}
ggplot(data = null_dist, aes(x = stat)) +
  geom_histogram()
```


6. How many of these `null` permutations have a difference of at least `obs_stat`?
none
```{r }
null_dist%>%
    filter(stat>=obs_diff) %>%
    tally()
```

Now that the test is initialized and the null distribution formed, you can calculate the p-value for your hypothesis test using the function `get_p_value`.

```{r inf-weight-habit-ht-pvalue}
null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")
```

This the standard workflow for performing hypothesis tests.

7.  Construct and record a confidence interval for the difference between the weights of those who exercise at least three times a week and those who don't, and interpret this interval in context of the data.
the confidence interval is not reflecting the value claimed by null hypothesis so we can reject the null hypothesis. There exists a significance difference in mean value of weights for the stated students group

```{r }
set.seed(1234)
null_dist %>%
  get_ci(level = 0.95)

yrbss %>% 
  filter(!(is.na(physical_3plus) | is.na(weight))) %>%
    specify(weight ~ physical_3plus) %>%
    generate(reps = 1000, type = "bootstrap") %>%
    calculate(stat = "diff in means", order = c('yes', 'no')) %>%
    get_ci(level = 0.95)
```
* * *

## More Practice

8.  Calculate a 95% confidence interval for the average height in meters (`height`) and interpret it in context.
```{r}

yrbss %>% 
    filter(!is.na(height)) %>% 
    specify(response = height) %>%
    generate(reps = 1000, type = 'bootstrap')  %>%
    calculate(stat = 'mean')  %>%
    get_ci(level = 0.95)

```

9.  Calculate a new confidence interval for the same parameter at the 90% confidence level. Comment on the width of this interval versus the one obtained in the previous exercise.
Since the p-value is less than 0.05 we reject the null hypothesis and conclude that there is a difference in average height of those who are physically active at least 3 days per week, and those who aren’t.
```{r}
yrbss %>% 
    filter(!is.na(height)) %>% 
    specify(response = height) %>%
    generate(reps = 1000, type = 'bootstrap')  %>%
    calculate(stat = 'mean')  %>%
    get_ci(level = 0.90)

```

10.  Conduct a hypothesis test evaluating whether the average height is different for those who exercise at least three times a week and those who don't.

u1 = average height of students who exercises at least 3 times a week 
u2 = average height of students who don't exercises at least 3 times a week:

H0: u1=u2 
HA: u1 != u2
```{r}
obs_diff <- yrbss %>%
  filter(!(is.na(physical_3plus) | is.na(height))) %>%
  specify(height ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
obs_diff
yrbss %>%
  filter(!(is.na(physical_3plus) | is.na(height))) %>%
  specify(height ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no")) %>% 
  get_p_value(obs_stat = obs_diff, direction = "two_sided")
```
Reject null hypothesis


11.  Now, a non-inference task: Determine the number of different options there are in the dataset for the `hours_tv_per_school_day` there are.
7
```{r}
(yrbss %>% 
   filter(!is.na(hours_tv_per_school_day)))$hours_tv_per_school_day %>%
  unique() %>%
    length()

```
12. Come up with a research question evaluating the relationship between height or weight and sleep. Formulate the question in a way that it can be answered using a hypothesis test and/or a confidence interval. Report the statistical results, and also provide an explanation in plain language. Be sure to check all assumptions, state your $\alpha$ level, and conclude in context.
```{r}
yrbss <- yrbss %>% 
  mutate(seven_hours_sleep = 
           ifelse(school_night_hours_sleep==7, 'yes', 'no'))
yrbss %>%
  filter(!(is.na(seven_hours_sleep) | is.na(weight))) %>%
    count(seven_hours_sleep)
```

he sample size is large enough to hold the central limit theorem. All the conditions for inference are met.


```{r}
obs_diff <- yrbss %>%
  filter(!(is.na(seven_hours_sleep) | is.na(weight))) %>%
  specify(weight ~ seven_hours_sleep) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
obs_diff

yrbss %>% 
  filter(!(is.na(seven_hours_sleep) | is.na(weight))) %>%
    specify(weight ~ seven_hours_sleep) %>%
    generate(reps = 1000, type = "bootstrap") %>%
    calculate(stat = "diff in means", order = c('yes', 'no')) %>%
    get_ci(level = 0.95)
```
Null hypothesis rejected

* * *
