---
title: "Eigenvalues and Eigenvectors"
author: "Keeno Glanville"
date: "2023-02-12"
output: html_document
---

```{r setup, include=FALSE}
library(pracma)
```

# Problem Set 1

## 1 What is the rank of the matrix A?

```{r}
A<- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), nrow = 4, ncol=4, byrow = TRUE)
A
# The rank of the matrix is determined by the number of number of independent rows. These are the non-zero rows in row echelon form

reA<-rref(A)
reA
#This matrix has rank 4
```




## 2. Given an mxn matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?
The maximum rank of a matrix is equal to the value of the lowest dimension of mxn



## 3. What is the rank of matrix B?
```{r}
B<- matrix(c(1,2,1,3,6,3,2,4,2), nrow=3, byrow= TRUE)
B

reB<- rref(B)
reB

#The rank of this matrix is 1
```



# Problem Set 2

Av − λv = (A − λI)v = 0
```{r}
A <- matrix(c(1, 2, 3, 0, 4, 5, 0, 0, 6), nrow=3, byrow=TRUE)
A
# characteristic polynomial of a matrix A is given by the determinant of A- lambda * Identity matrix 
# For this example we will be using lambda to be x

A <- matrix(c('1-λ', 2, 3, 0, '4-λ', 5, 0, 0, '6-λ'), nrow=3, byrow=TRUE)
A

```
 determinant of A : (1-λ)((4-λ)(6-λ) - 5*0) - 2*0*0 = (1-λ)((4-λ)(6-λ))
 characteristic equation: -λ^3 + 10λ^2 - 24λ + 24
 lambda values (1-λ)((4-λ)(6-λ)) = 0
### eigeinvalues of matrix λ = 1, 4, 6


### eigen vector
To find the eigen vector we know:  Av − λv = (A − λI)v = 0

we now have find the eigen vectors

### 1. First eigenvector
```{r}
λ = 1
A1 <- matrix(c(1-λ, 2, 3, 0, 4-λ, 5, 0, 0, 6-λ), nrow=3, byrow=TRUE)
A1

# Thus we need a vector v that creates proves the statement (A - λI)v = 0

```
0v.1 +2v.2 +3v.3 = 0 
0v.1 +3v.2 +5v.3 = 0
0v.1 +0v.2 +5v.3 = 0

V1 = 0, V2 = V3



*Eigenvector for λ = 1 : (0, 0, 1)*






### 2. Second Eigenvector
```{r}
λ = 4
A1 <- matrix(c(1-λ, 2, 3, 0, 4-λ, 5, 0, 0, 6-λ), nrow=3, byrow=TRUE)
A1

# Thus we need a vector v that creates proves the statement (A - λI)v = 0

```
-3v.1 +2v.2 +3v.3 = 0 
0v.1 +0v.2 +5v.3 = 0
0v.1 +0v.2 +2v.3 = 0

V1 = 0, V2 = 1, V3 = 0



*Eigenvector for λ = 4 : (0, 1, 0)*






### 3. Third Eigen vector
```{r}
λ = 6
A1 <- matrix(c(1-λ, 2, 3, 0, 4-λ, 5, 0, 0, 6-λ), nrow=3, byrow=TRUE)
A1

# Thus we need a vector v that creates proves the statement (A - λI)v = 0

```
-5v.1 +2v.2 +3v.3 = 0 
0v.1 −2v.2 +5v.3 = 0
0v.1 +0v.2 +0v.3 = 0

V1 = 0, V2 = 0, V3 = 1



*Eigenvector for λ = 6  (0, 0, 1)*
















